\documentclass{report}
\begin{document}
\title{STAPL Alpha Release Mini\-Reference -- VERSION 0.5}
\date{\today}
\maketitle

\newenvironment{hashitemize}{%
  \renewcommand{\labelitemi}{\bfseries \#\#}%
        \begin{itemize}}{\end{itemize}}

\newcommand{\stapl}{{\sc STAPL}}
\newcommand{\stl}{{\sc STL}}
\newcommand{\pview}{{\sc PVIEW}}
\newcommand{\pviews}{{\sc PVIEWS}}
\newcommand{\pcontainer}{{\sc PCONTAINER}}
\newcommand{\pcontainers}{{\sc PCONTAINERS}}

\pagestyle{myheadings}
\renewcommand{\chaptermark}[1]{
              \markboth{*** DRAFT 0.5 *** \chaptername\ \thechapter. #1 } {} }
\renewcommand{\sectionmark}[1]{
              \markright{*** DRAFT 0.5 *** \thesection. #1 } {}}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\vspace*{8cm}
\begin{center}
Copyright (c) 2014, Texas Engineering Experiment Station (TEES), a
component of the Texas A\&M University System.

All rights reserved.
\end{center}
\pagebreak

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\chapter{Introduction}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\section{What is \stapl?}

\stapl\ is a framework for parallel C++ code development.

What is a \stapl\ program?  It has three parts:
\begin{itemize}
\item Dependence graph
\newline
The dependence graph describes which memory operations must precede others,
and which computations produce values that others consume.
\item Data distribution
\newline
The data distribution describes how data aggregates should be partitioned
among the memories associated with processing elements.
\item Instructions
\newline
The instructions, as in any program, describe the computations.
\end{itemize}

The instructions are written in C++, using templated containers and algorithms
analogous to those in the Standard Template Library (\stl\ ).
The dependence graph is defined by the instructions written in C++.
The data distribution takes on default values
in the absence of an explicit specification.

The \stapl\ system has the following characteristics:

\begin{itemize}

\item Execution Model

The \stapl\ execution model is an asynchronous coarse-grain data flow machine.
\stapl\ enables the programmer to describe the parallelism in the program
at the finest granularity possible.  The \stapl\ system will coarsen the grain
of the described parallelism to fit the execution platform,
both at compile time and at execution time.

\item Memory Model

The \stapl\ memory model is a global address space with automatic data
distribution.  This is often referred to as Partitioned Global
Address Space (PGAS).  Data is never replicated in the memories associated
with processing elements.

\item Extensibility
\newline
The system is extensible because new kinds of each of the major features can
be introduced by composition, inheritance, and even implementation from scratch.
Programmers can create new kinds of views, containers, distributions,
algorithms, schedulers, and skeletons.

\item Engineering Excellence
\newline
The implementation of \stapl\ is engineered to be usable for real-world problems
on the full range of parallel systems, providing robust, high-performance
execution.
\end{itemize}

The following features implement the design principles of \stapl.

\begin{itemize}

\item Execution Model
\begin{itemize}
\item Multi-protocol parallelism:
\newline
Applications built with \stapl\ can employ message-passing on a coarse-grain
basis, and threading on a fine-grain basis, without any effort required
from the programmer.
\end{itemize}

\item Memory Model
\begin{itemize}
\item Transparent data distribution:
Programmers can create applications using a shared memory address space and
rely on \stapl\ to manage the distribution and data movement across distributed
memory configurations.  In addition to the default behavior, the programmer
can also select from a library of data distribution mechanisms, or even write
their own.
\newline
\end{itemize}

\item Extensibility
\begin{itemize}
\item Nested parallelism:
\newline
Programmers can compose parallelism using generic containers
and generic algorithms.
\item New container classes:
\newline
Programmers can use parallel data structures commonly employed in High Performance Computing which do not have serial analogs in the \stl\
%% FOR THE BETA RELEASE
%\item Library Interoperability
%\newline
%Applications built with \stapl\ will run in parallel with numerical libraries commonly used in High Performance Computing, such as LAPACK.
\end{itemize}

\item Engineering Excellence
\begin{itemize}
\item Scalable performance:
\newline
Applications built with \stapl\ will run in parallel on systems from
multi-processor tablets to the largest supercomputers.
\item Platform independence:
\newline
Applications built with \stapl\ can be deployed on both shared memory and distributed memory systems, requiring only a re-compile and re-link.  Execution is virtualized, providing a runtime system that is independent of the number or type of processors.
\end{itemize}

\end{itemize}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\chapter{A Complete \stapl\ Example Program}

This chapter presents a complete \stapl\ program, in order to motivate
the material that follows.  It reads numeric input from a file specified
on the command line, computes some simple descriptive statistics, and
displays the results to the user.  Comments in the code indicate the
name of the chapter in which the corresponding construct is discussed.

\vspace{0.2cm}
\noindent
\rule{12cm}{0.5mm}
\vspace{0.2cm}

\input{stats}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\chapter{C++ Concepts for Parallel Programmers}

This chapter briefly describes C++ concepts and features that are
essential for understanding the remainder of this document.

\section{Class Declarations}

A struct is a class in which the members are default public.
Since it is undesirable to have data members which have public access,
we adopt the convention in this document
that classes with no data members are defined using the
\texttt{{\bf struct}}
keyword rather than the
\texttt{{\bf class}}
keyword.

\section{Function Objects}

A function object is a class that defines
\texttt{{\bf operator()}}.
This operator is known as the "function call" or "application" operator.

Besides the application operator, a function object can also have
data members.  This enables the function to have state.

Each function object has its own type, unlike ordinary functions, which
have different types only when they have different signatures.

Because a function object is an object, it can be passed as a argument.
An alternate name for function object is "functor".

C++11 also supports lambdas, which are not used in this document.

\section{Templates}

A template is a class or function that is specified in terms of types
or values, which are treated as parameters of the template.

\section{Standard Template Library}

The Standard Template Library (\stl) is a part of the C++ standard library.
It provides the programmer with the ability to use \emph{generic programming}.
The \stl\ includes collection classes for organizing data, and algorithms
for applying to those collections.  The components of the \stl\ are all
template, so they can be used with any elemental type.

The \stl\ has three main features:
\begin{itemize}
\item
Containers
\newline
Containers provide varying interfaces to collections of objects.
These interfaces implement common Abstract Data Types.
\item
Iterators
\newline
Iterators are used to access the elements in collections of objects.
The chief benefit of iterators is that they provide a simple and common
interface for all of the container types.
\item
Algorithms
\newline
Algorithms use iterators to process the elements of containers.
\end{itemize}

\section{Namespaces}

In the examples in this text, all \stapl\ constructs are qualified with
\texttt{{\bf stapl::}}
to emphasize the \stapl\ functionality, rather than use a
\stapl\ namespace.
This also avoids conflicts with \stl\ entities.  We recommend that users
follow the same convention.  The alternative is to qualify all
constructs with
\texttt{{\bf std::}}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\chapter{Parallelism Concepts for C++ Programmers}

This chapter briefly describes parallelism concepts that are
essential for understanding the remainder of this document.

\section{SPMD}

SPMD (Single Program, Multiple Data)
is a technique employed to achieve parallelism.
In SPMD, multiple autonomous processors simultaneously execute
the same program at independent points.

\section{MPI}

Message Passing Interface (MPI) is a de facto standard that supports portable
message-passing programming in Fortran, C and and C++. It consists of a set
of library routines that provide protocols for point-to-point and collective
communication. Implementations are available on common processor architectures and operating systems.

\section{OpenMP}

Open Multi-Processing (OpenMP) is a standard that supports multi-platform
shared memory multiprocessing programming in Fortran, C, and C++.
It consists of a set of compiler directives (Fortran) or pragmas (C/C++), library routines, and environment variables that influence run-time behavior.
Implementations are available on common processor architectures and operating systems.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\chapter{Relationship between \stapl\ and \stl\ }

\stapl\ has functionality similar to the \stl.  \stapl\ executes on uni-
or multi-processor architectures with shared or distributed memory and
can co-exist in the same program with \stl.  \stapl\ top-level components
have well-defined counterparts in \stl.

\begin{itemize}
\item
pContainers - parallel equivalents of \stl\ \emph{containers}.
\item
pAlgorithms - parallel equivalents of \stl\ \emph{algorithms}.
\item
pViews - provide data access operations independent of the underlying
storage, as do \stl\ iterators.
\end{itemize}

Having a basic understanding of the major features of \stl\ will make it
much easier to learn \stapl.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\chapter{\stapl\ Concepts}

This chapter briefly describes \stapl\ concepts that are
essential for understanding the remainder of this document.

\section{Parallel Containers}

\stapl\ parallel containers (\pcontainers) are distributed,
thread-safe, concurrent objects.  The are composable and extensible
by C++ inheritance mechanisms.  Most \pcontainers\ have analogs in the
\stl.  There are several additional containers not provided by the \stl\
which are very important for High Performance Computing: the parallel
multi-dimensional array (pMultiarray) and the parallel graph (pGraph).

\pcontainers\ provide methods that correspond to the \stl\ container
methods, where there is an analogous container.  They also provide
additional methods specifically designed for parallel use.

While the data in a \pcontainer\ is distributed, \pcontainers\ provide a
\emph{shared object view}.  This means that they are shared data structures
with a global address space.  Each \pcontainer\ element has a unique
global identifier (GID).  \stapl\ automatically translates the GID so that
it can locate both local and remote elements.

The following containers are supported by \stapl:
\begin{itemize}
\item
Sequence containers: {\tt list, vector }
\item
Associative containers: {\tt set, map }
\item
Indexed containers: {\tt array, matrix, multiarray (vector is also indexed) }
\item
Relational containers: {\tt graph, dynamic\_graph, multi\_graph }
\end{itemize}

\section{Iterators, Ranges, and Views}

The \emph{iterator} is a common abstraction used to represent data
in sequential programming.
Iterators are used in generic algorithms, such as
the \stl\ in C++, to specify the data.
A key function of the iterator is to separate data structures from algorithms.
All accesses to data are performed through iterators, which are implemented
as memory pointers.  Iterators cannot encapsulate other information about the
referenced data, e.g., length of the sequence.
Moreover, the operations associated with iterators are essentially
limited to {\em sequential} traversals of a collection of elements,
and so algorithms using iterators have a linear, flat memory model
that only supports {\em sequential access}.

To allow concurrent access to multiple elements,
some parallel systems provide the \emph{range}.
A range defines a sequence of elements between two
iterators. With the range abstraction, the program can create
sub-ranges and process them in parallel.  Ranges alleviate the
sequential access problem, but still offer only a flat memory model.

In \stapl, a \pview\ encapsulates the information about the collection
of elements it represents, the elements that can be accessed through it,
and provides a set of operations to interact with the data.
Using \pviews, parallel algorithms can be specified using abstract data
types (ADTs).  The \pview\ encodes information about the partition of
the data that allows it to operate in a distributed environment.
The \emph{domain} of a \pview\ is the valid set of identifiers.
The domain specifies which elements of the underlying container
can be accessed through the \pview.

\section{Parallel Views}

Different ADTs provide common sets of operations or concepts.
Views can have any of the following attributes, which control the
operations they can perform.

\texttt{read\_write}:
Views that have this attribute support operations to read and
write values for a given element identifier (i.e., index or key).
Some views have the \emph{read} attribute, but not the \emph{write}
attribute.

\texttt{subscript}:
Views that have this attribute support the square-bracket operator.
This operator returns a reference over the element instead of
returning a copy of the value, as the
\texttt{read}
operation does.

\texttt{sequence}:
Views that have this attribute support the operations
\texttt{begin},
and
\texttt{end}
and the required associated types
required for use with iterators.

\texttt{insert\_erase}:
Views that have this attribute provide the dynamic
operations \emph{insert} and \emph{erase}.

\stapl\ views fall into the following categories:

\begin{itemize}
\item
Model abstract data types:
\newline
array, vector, list, map, multiarray, heap, set, graph
\item
Modify access patterns to the container:
\newline
repeated, overlap, counting, strided, reverse, filter, transform, linear
\item
Modify distribution of the container:
\newline
native, balance, explicit, segmented
\end{itemize}

\section{Work Functions}

A work function is a function object which is applied to each element
of a view by a basic algorithm (see below).
\texttt{{\bf operator()}}
specified by the work function
should have as many arguments as there views passed to the basic algorithm.

A work function can have other members besides
\texttt{{\bf operator()}}.
It can have data members, which hold state.
If the function object has data members, the class will need
a constructor and/or "setter" method to give them values.  It will
also need a
\texttt{{\bf define\_type}}
method (see Chapter 3).
It may also need a "getter" method to retrieve updated state.

\section{Basic Algorithms}

These functions take one or more work functions as arguments,
which are applied to the elements of the specified view(s),
according to the dependence graph which defines the semantics
of the particular basic algorithm.

\texttt{{\bf map\_func }}

When processing a single view, apply the work function to each
item of the view independently.  When processing multiple views,
apply the work function to a corresponding value from each view.
The work function may return a void type, if it will not be used
to create nested parallelism.

The following pseudocode shows how map\_func acts on simple vectors.

\begin{verbatim}
  p <- [2 3 5 7 11]
  f <- [1 1 2 3 5 ]
  map_func( plus, p, f )
3 4 7 10 16
\end{verbatim}

\texttt{{\bf map\_reduce }}

Apply the first work function independently to each item of the view.
Apply the second work function cumulatively to each item of the values from
the first application.

This function is sometimes used with an identity function for
the map operation, thus becoming a simple reduction.
Both of the work functions must return a non-void type.

\begin{verbatim}
  p <- [2 3 5 7 11]
  f <- [1 1 2 3 5 ]
  map_reduce( neg, plus, p, f )
-3 -4 -7 -10 -16
\end{verbatim}

\texttt{{\bf reduce }}

Apply the first work function independently to each item of the view.
The work function must return a non-void type.

\begin{verbatim}
  p <- [2 3 5 7 11]
  reduce( plus, p ]
28
\end{verbatim}

\texttt{{\bf scan }}

Apply the work function cumulatively to each item of the view,
saving the intermediate results of each function application.
The work function must return a non-void type.

\begin{verbatim}
  f <- [1 1 2 3 5 ]
  scan( plus, f )
1 2 4 7 12
\end{verbatim}

\texttt{{\bf serial }}

Apply the work function to each item of the view in serial order.
The work function may return a void type, if it will not be used
in a nested fashion.

\texttt{{\bf serial\_io }}

Apply the work function to each item of the view in serial order,
and execute it on location zero.
The work function may return a void type, if it will not be used
in a nested fashion.

\texttt{{\bf do\_once }}

Execute the work function a single time on location zero.

All \stapl\ constructs are qualified with
\texttt{{\bf stapl::}}
to emphasize the \stapl\ functionality, rather than use a
\stapl\ namespace.
This also avoids conflicts with \stl\ entities.  We recommend that users
follow the same convention.  The alternative is to qualify all
constructs with
\texttt{{\bf std::}}.

\section{Nested Parallelism}

\stapl\ implements parallel algorithms and data structures using standard C++,
which provide an SPMD model of parallel, and which support nested (recursive)
parallelism.  Nested parallelism is important because:

\begin{itemize}
\item
Many large parallel systems have a hierarchical architecture onto which
nested parallelism would naturally map.
\item
Library functions are often used as basic building blocks which themselves
can be nested or incorporated into a larger parallel application.
\end{itemize}

\section{Portability}

\stapl\ provides portability across multiple systems by including its own
runtime system.  This supports high level parallel constructs (e.g. task
graph execution) and their low level implementation.  \stapl\ also provides
its own scheduling and data distribution interfaces.  So, there is no need
to modify user code when porting a \stapl\ application from one system to
another.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\input{cont}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\input{view}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\input{alg}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\input{io}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\end{document}
